
# Chapter 3: Running the statistical techniques

- Run statistics on data
    - Exercises using different techniques
    - CI vs p-value... non-significant but still important
- Extract results or numbers from model
    - Using broom

### Exercise 3: A simple logistic regression
NE: Run logistic regression on data. Get relevant information from the model.
Logistic regression is often used when studying how factors associate with disease
(since disease state is usually binary: they have it or they don't)


```{r}
load("datasets/nickel.rda")
glm()
```

### Exercise 4:
NE: Run Mixed effects models on data. Get relevant information from the model. 

### Exercise:
- MCQ: here is a research question, here is the result of an analysis
(non-significant but wide CI that is protective). which is more appropriate
interpretation of the result:
    - There is no association.
    - while non-significant, this effect 

## Lesson 2: Logistic regression, OR, and RR
### Exercise 1: NE Calculate OR (simulated data) and RR. Compare.

```{r}
```

### Exercise 2: MCQ Interpretation of OR vs RR vs probability.
### Exercise 3: MCQ Multiple logistic regression interpretation
