
# Chapter 2: Data wrangling and formatting

- Wrangling data to get into proper for statistics
- Transforming for statistics if necessary

## Lesson 1: Here's the data ... tidy it up!
### Exercise 1: Which are untidy data?
Identify tidy datasets given the statistical technique/study design
(or which are wrong)

- wrong ones (possibles):
    - wide instead of long for longitudinal study
    - wide instead of long for those with the disease and those without
    - different character strings for same disease (rather than numbers), e.g.
    "t2dm" and "dm" for diabetes status
    - "excel" style of two datasets in one spreadsheet
    - 

### Exercise 2: So then, tidy it up!
NE:

... maybe two waves that need to be combined?

```{r}

```

### Exercise 3: Long vs wide, or which side is which?
NE: Sometimes analyses are more easily done with data all completely in a long
form, either for plotting or for summarizing.

```{r}

```

### Exercise 4: Data could be tidy, but not for analysis

NE: We have a longitudinal cohort that has some variables 


```{r}

```
