```{r setup}
source("preamble.R")
```

# Chapter 3: Running the statistical techniques

**Objectives:**

- Learn to run some common statistic techniques on different datasets and research
questions
- Learn how to extract the relevant results from these stats methods
- Understand why its important to not use or even look at p-values in the output
- Know how to understand and interpret the results (we'll get to knowing what
exactly is most useful to present and show for higher impact)
- Importance of examining multiple levels of adjustment (unadjusted, minimally,
fully), and how this can inform what is going on in the relationship.
- Knowing how to choose/decide on what variables to adjust for

## Lesson 1: Common statistical techniques used for analyzing cohorts

*Content*:

- Common statistical techniques:
    - mixed effects modeling
    - GEE (?)
    - cox proportional hazard models
    - linear regression
    - logistic regression
    - Poisson regression
    - Conditional logistic regression (for nested case-cohort designs)

*Exercises*:

- NE: Change variables scaling or transform them to see how the estimates change...
what does that mean for interpretation? (or lesson 4)

### Exercise: Type of statistical method is important, especially for longitudinal studies

Before the development of mixed effects modeling, analyzing longitudinal data
was fairly difficult. This was because the early methods, which typically were 
some type of ANOVA or simple regression, had an assumption of independence in the
data. But longitudinal data is not independent... A single individual can have 
multiple measures done over time and are thus dependent.

Take a look at the difference between using a logistic regression model compared
to a mixed effect model.

```{r}
# TODO: Check to confirm exponentiating of output.
logistic_model <- glm(prevchd ~ totchol + period, data = framingham, family = binomial)
summary(logistic_model)$coef

# show this in lesson 4?
# tidy(logistic_model, conf.int = TRUE, exponentiate = TRUE) %>% 
#     select(term, estimate, conf.low, conf.high)

mixed_model <- lmer(prevchd ~ totchol + period + (1 | randid), data = framingham)
summary(mixed_model)$coef

# show this in lesson 4?
# tidy(mixed_model, conf.int = TRUE) %>% 
#     select(term, estimate, conf.low, conf.high)
```

## Lesson 2: Adjustment, confounding, and modelling.

*Content*:

- What it means to "adjust" for confounders? What is a confounder? (probably explained 
in one of the other epi courses, so be brief here).
    - What does adjustment mean when time is included? Gets trickier. This is why
    cross-sectional analyses are simpler analytically.
    - Danger of not controlling for variables: Simpson's Paradox. Can lead to actual
    harm if analyses not thought through and properly analyzed.
    - Techniques to determine what to adjust for (DAG, literature, IC)
    - Model selection (AIC, BIC, Cross-validation; when to use each)

*Exercises*:

- MCQ: Which options would be most appropriate for using various model selection techniques?
    - list different conditions with possible methods.
- NE: Cross-validation of a model.
- MCQ/text: Describe different impact of adjusting, etc. including collider variables.
How does that change the interpretation? How can this influence the health impacts
if inappropriately published? (or in lesson 4)

### Exercise: Understanding confounding pathways

{{Probably as a Bullet/TabExercise}}

- MCQ/text: Present a DAG of hypothesized variables and pathways. Which of the
follow (or write out which) are the important variables to consider/adjust for.

```{r simple_confounding}
grViz("
digraph {
    node [style = filled fillcolor = none]
    Exposure [fillcolor = 'LightBlue'] 
    Outcome [fillcolor = 'OrangeRed']
    Exposure -> Outcome
    Confounder -> {Exposure Outcome}
}")
```

This simple confounding example shows how each variable relates to each other in
hypothesized pathways. Understanding how variables confounded the relationship
between the exposure and the outcome is essential to drawing more accurate
inferences about the associations. Creating these DAGs of the hypothesized
pathways is a powerful tool to understanding what could be confounders, what
could be colliders, and what needs to be adjusted for. Which are more
appropriate choices:

```{r confouding_example}
grViz("
digraph {
    node [style = filled fillcolor = none]
    BodyFat [fillcolor = 'LightBlue'] 
    CVD [fillcolor = 'OrangeRed']
    BodyFat -> CVD
    Sex -> {BodyFat CVD Testosterone}
    Testosterone -> {BodyFat CVD}
}")
```

- Options: {{likely all answers will be correct.. or maybe rank them ...}}
    - Since sex influences testosterone and since both sex and testosterone
    are confounders, only need to adjust for either of these variables to control
    for the confounding pathway.
    - Adjust for both

```{r collider_bias}
grViz("
digraph {
    node [style = filled fillcolor = none]
    BodyFat [fillcolor = 'LightBlue'] 
    CVD [fillcolor = 'OrangeRed']
    BodyFat -> CVD
    Sex -> {BodyFat CVD Testosterone ExerciseType}
    Testosterone -> {BodyFat CVD}
    ExerciseType -> {BodyFat CVD}
}")
```


### Exercise

- NE: When you are unsure of which variables may provide a better fit and maybe
explain the model and results more, using model information criterion methods can
help narrow the possible models down. For instance, when deciding on appropriate
covariates to adjust for. Run this command to identify which model is the "best"
of those compared.

```{r}
# TODO: Make more appropriate models (more typically seen in real analyses).
# TODO: Confirm that MuMIn is the best package for learners to use.
library(MuMIn)
m1 <- glm(cvd ~ totchol, data = framingham, family = binomial)
m2 <- glm(cvd ~ totchol + sex, data = framingham, family = binomial)
m3 <- glm(cvd ~ totchol + sex + bmi, data = framingham, family = binomial)

model.sel(m1, m2, m3)
```

### Exercise

NE: Run model without adjusting, with adjusting, and lastly with adjusting for an
inappropriate variable (how that changes things)

```{r}
# TODO: Confirm that period and randid are sorted properly.
# TODO: Confirm that cvd variable is the right one (from raw data).
# TODO: Check that tidy exponentiation works for this method.
unadjusted <- lmer(prevchd ~ totchol + period + (1 | randid), data = framingham)
tidy(unadjusted, conf.int = TRUE)

adjusted <- lmer(prevchd ~ totchol + sex + cursmoke + period + (1 | randid), data = framingham)
tidy(adjusted, conf.int = TRUE)

inappropriate <- lmer(prevchd ~ totchol + sex + cursmoke + age + period + (1 | randid), data = framingham)
tidy(inappropriate, conf.int = TRUE)
```

### Exercise: Followup MCQ/text to above

Given the results, what are some possible conclusions/interpretations?
How might the adjustment for age AND period influence the results?

## Lesson 3: Interaction testing and sensitivity analyses

*Content*:

- Sensitivity analyses
    - exploring potential sub-hypotheses to test assumptions of models
    - e.g. stratified analysis, models with and without covars to id which is
    impacting results, etc
- Simpson's paradox? need to consider subgroup analyses (interactions), especially
for sex and ethnicity. (Expand more here, tho mentioned in lesson 2)
- Interaction testing, especially for sex and ethnicity (common requirement in order
to publish)

*Exercises*:

- NE: Coding example of Simpson's paradox in health setting (real world example?) (or in lesson 2)
- NE: Maybe exercise where some observations are removed and show how that impacts
the p-value...? to highlight how unreliable it is?

### Exercise

- NE: Interaction testing for prospective study. Interpretation


## Lesson 4: Extracting relevant data from results and post-modelling-wrangling {{wording needs changes}}

*Content*:

- Using broom, what to take from broom, conf.int
- Interpreting OR, RR, IRR (incidence risk ratio), and other forms of estimates
- Quick overview of the unreliability of p-values, the danger of using them to
inform clinical and public health policy (find examples?)

*Exercises*:

- NE: Example of how small changes to data can influence p-value.
- MCQ/text: Which are the most appropriate interpretations for OR, RR, IRR?
- MCQ: here is a research question, here is the result of an analysis
(non-significant but wide CI that is protective). which is more appropriate
interpretation of the result:
    - There is no association.
    - while non-significant, this effect 

# Notes:
- Statistics (more of a review, expect them to know what they are doing):
    - Logistic regression
    - Mixed effects modelling
- Choice of statistic is dependent on question asked.
- In general, cohorts try to address questions such as:
    - "what type of exposures increase the risk of disease?"
    - "how much and how long do individuals need to be exposured to a risk factor
    to develop the disease?"
    - "For those that have a disease, how do they differ from those without?"
    - "Those that have more exposure over time, are they more likely to develop a disease?"
- These questions generally require some type of regression modelling in order to
estimate magnitude of association and the uncertainty around that association.
